Blocks of extended ancestry around selected loci in hybrid zones

### 2016-02-17 drafts/hybrid_zone.tex
edits to intro

Because of rapid removal from the population, we expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.

The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.  
These models predict cline width to be narrowest \yb{(i.e. the transition in allele frequency is the steepest)} at loci close to selected loci, and cline width at less tightly linked loci to be wider. %and more heavily influenced by demography.} %hybrid zone age and rates of gene flow. 
\yb{The focus on cline width provides an expectation for how selection shapes covariation in ancestry at a single locus over \emph{geographic space}, and our aim is to extend this to obtain an expectation for how selection shapes covariance in ancestry over \emph{genomic space}}

@@

Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the strength of selection at linked loci. 
%Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.   % YB I don't think we are actually asking about density of selected sites, are we? 
Our goal here is  to understand how the strength of \yb{selection} and hybrid zone age interact to produce observed patterns of ancestry \yb{around selected sites in recently established hybrid zones}. %in present day populations.  
This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness. 
In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
If hybrids have non-zero fitness, and some amount of introgression can occur, patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}. 
\yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci \citep{Porter1997, Gompert2012}, but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}  
%In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation. 
%In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.

Our naive expectation under a basic underdominance model is that, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center. 
This is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events. 
Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.  
\yb{The theoretical expectation developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}

\alisa{Moved some points here to the discussion.}

>>>>

The extensive history of population genetic modeling of hybrid zones has focused largely on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}.  
These models predict cline width to be narrowest \yb{(i.e. the transition in allele frequency is the steepest)} at loci close to selected loci, and cline width at less tightly linked loci to be wider. %and more heavily influenced by demography.} %hybrid zone age and rates of gene flow. 
\yb{This provides some expectation for how selection shapes covariation in ancestry at a single locus over \emph{geographic space}, and our aim here  is to extend this to include expectations for how selection shapes covariance in ancestry over \emph{genomic space}}.

Covariance in ancestry across a genome provides information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the relative strength of selection affecting subsets of the genome. Our goal is to understand how selection against hybrids interacts with dispersal and recombination interact to produce observed patterns ancestry around selected sites in recently established hybrid zones. In addition to interpreting non-equilibrium patterns of diversity in hybrid zones, this could be especially important for understanding the impact of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness. In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
If, on the other hand, hybrids have non-zero fitness, some amount of introgression can occur, and  patterns of introgression can provide information about selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}. 
\yb{In this latter case, heterogeneous patterns of ancestry across the genome potentially allows researchers to identify putative targets of selection in hybridizing populations.}  
\yb{To date, researchers have done so by examining heterogeneous clines in ancestry across loci \citep{Porter1997, Gompert2012}, but have not made use of the information contained in the length of ancestry blocks around putatively selected loci.}

Under a model of selection against hybrids, migrant haplotypes containing the selected locus will be removed rapidly from the population, preventing introgression in surrounding genomic regions. We therefore expect to observe \yb{a deficit of} short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone. As a corollary to this , conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center.  This is because an unfit haplotype is more likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.

Using a combination of predictions from theory and simulated hybrid zones, we present a model of genome-wide patterns coancestry as they relate to hybrid zone age, genetic distance from selected locus and geographic distance from hybrid zone center.  
\yb{The theoretical expectation developed in this manuscript can aid in the identification of loci under selection in hybridizing populations.}

@@

%Covariance in ancestry across a genome can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone and the strength of selection at linked loci. 
%Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations.   % YB I don't think we are actually asking about density of selected sites, are we? 
%Our goal here is  to understand how the strength of \yb{selection} and hybrid zone age interact to produce observed patterns of ancestry \yb{around selected sites in recently established hybrid zones}. %in present day populations.  
%This may be especially important for our understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctness. 
%In the extreme case of zero fitness in first-generation hybrids, for example, parental genomes will never mix and the stable hybrid zone will resemble a recently formed one. 
%If hybrids have non-zero fitness, and some amount of introgression can occur, patterns of introgression can provide information about the strength of selection \yb{in addition to the} extent of hybridization, and timing of secondary contact \citep{Price2009, Hellenthal2014}.

%In this latter case, it is also of interest to identify putative targets of selection, as these will provide insights into the biological mechanisms of speciation. 
%In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state.

%Our naive expectation under a basic underdominance model is that, conditional on having the ancestry that is at lower frequency (that is, being on the `wrong' side of the hybrid zone), the length of unbroken ancestry surrounding the selected locus will be relatively long when far away from the zone center. 
% is because an unfit haplotype is likely to be a recent migrant from the other side of the hybrid zone, and therefore will not have experienced many recombination events.
